---
title: "Siane"
author: "Nuno Carvalho dos Santos"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Siane}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---


## Introduction

[Siane](https://github.com/Nuniemsis/Siane/) (El Sistema de Información del Atlas Nacional de España) is a project that supports technologically the publications and productions of the National Spanish Atlas(ANE). Recently, this project released CARTOSIANE, a set of maps compatible with the National Institute of Statistics georreferenced data(INE).
The aim of this package is to create useful functions that plot INE's georreferenced data on Siane map's polygons.  

Siane collects maps from five different scales(1:3M, 1:6.5M, 1:10M, 1:14M, 1:60M). The scope of this project covers the first two scales, i.e, 1:3M and 1:6.5M. The remaining scales represent the countries globally. We are not using those scales. This project focuses only in spanish maps.  

Each set of maps is downloaded individually. [The README.md](https://github.com/Nuniemsis/Siane/) file has very detailed information about the downloading process.

```
SIANE_CARTO_BASE_E_14M.ZIP  
SIANE_CARTO_BASE_S_10M.ZIP  	  
SIANE_CARTO_BASE_S_3M.ZIP
SIANE_CARTO_BASE_S_6M5.ZIP
SIANE_CARTO_BASE_W_60M.ZIP  
```

This package needs SIANE_CARTO_BASE_S_3M.ZIP and SIANE_CARTO_BASE_S_6M5.ZIP files.

In this document I will describe and make use of the main functions of this package. The first step is to load the package. Uncomment the first two  code lines in case you haven't installed the package yet.  

```{r, warning = FALSE, message = FALSE , eval = FALSE}
#library(devtools)
#install_github("Nuniemsis/Siane")
library(Siane)
library(pxR)
library(RColorBrewer)

```

## Setting the path of the maps

As I already explained, Siane consists of a collection of maps. This package needs to locate the path of this collection(folder) in order to search the requested map.

```{r, eval = FALSE}
obj <- register_siane("/home/ncarvalho/Descargas/")
```

## Loading a map

Once we have located all the maps we can select one according with our data.
These parameters are enough to specify a map.  
  - `year` : Maps can change over time. This numerical parameter is the year of the map we want.  
  - `canarias` : It indicates whether we want to plot Canarias or not.
  - `level` : It's the administrative level. For this set of maps there are three: "Municipios", "Provincias" and "Comunidades"  
  - `scale` : The scale of the maps. The default scale for municipalities is 1:3000000 `scale <- "3m" `. For provinces and regions the default scale is 1:6500000 `scale <- "6m"`.  
  - `peninsula`: It's the relative position of the Canarias island to the peninsulae.  

```{r, eval = FALSE}
level <- "Municipios"
canarias <- TRUE
year <- 2016
scale <- "3m"
peninsula <- "close"
```

Now we call the `siane_map` function to extract a map from all the map's collection.

```{r, eval = FALSE}
shp <- siane_map(level = level, obj = obj, canarias = canarias, year = year, scale = scale, peninsula = peninsula) # Reading the map from the maps collection
```

## Plot the map

```{r,fig.show='hold', fig.width = 5, fig.height = 5, message=FALSE, eval = FALSE}
raster::plot(shp)
```

This function also allows the user to change the position of Canarias islands through the peninsula parameter. 

```{r, eval = FALSE}

peninsula <- "far"

shp <- siane_map(level = level, obj = obj, canarias = canarias, year = year, scale = scale, peninsula = peninsula) # Reading the map from the maps collection
```


```{r,fig.show='hold', fig.width = 5, fig.height = 5, message=FALSE, eval = FALSE}
raster::plot(shp)
```



The spanish political map looks just like this.  Let's try to plot some data on it. To keep following this tutorial you should download previously the data from the INE website. The README.md file explains how to download it. I will also share the [link](http://www.ine.es/jaxiT3/Tabla.htm?t=2879) here in case you already now how to do it.

## Polygon colour plot


#### Data frame reading 

Let's explore the dataset before plotting the data
```{r, eval = FALSE}
df <- as.data.frame(read.px("/home/ncarvalho/Descargas/2879.px"))
names(df) # List the column names  of the data frame
```

#### Data frame filtering 

First we need to understand the data frame. It's columns are the following:  
- `Periodo` is the time column in year's format.  
- `value` is a column with the numeric value of the population.  
- `Sexo` is the sex of that population.  
- `Municipios` is a character array with the municipality name and the municipality code
In this dataset there is only one value per territory.  

Split the `Municipios` column to get the municipality's codes.  
We are storing these codes in the column `codes`. This column is really important in order to plot the polygons with the corresponding colour colours. 


#### Data frame preparation 

```{r, eval = FALSE}
by <- "codes"
```

We create a single column in the data frame with those codes.
```{r, eval = FALSE}
df[[by]] <- sapply(df$Municipios,
                   function(x) strsplit(x = as.character(x), split = " ")[[1]][1])
```

```{r, eval = FALSE}
df$Periodo <- as.numeric(as.character(df$Periodo)) # Convert factor to numeric
```

Plotting polygons by colour intensity requires one unique value per territory.
Therefore, we have to filter the data frame.  
__Remember__: One value per territory.  
In this example I want to plot the total population in the year 2016.   
```{r, eval = FALSE}
df <- df[df$Sex == "Total" & df$Periodo == year, ]
```

The `siane_merge` function assigns each polygon a value. Those values are stored in a column of `shp_merged@data` which name can be chosen by the user changing the `value` variable.

```{r, eval = FALSE}
by <- "codes" # name for the codes column
value <- "value" # name for the values column

shp_merged <- siane_merge(shp = shp, df = df,by = by, level = level, value = value)

```


We can plot the map with the statistical data with the `plot` function.
The `RColorBrewer` package provides lots of colour scales. We can use this colour scales to visualize data over the map.
The following function displays all the colour scales from that package.


```{r, fig.width = 8, fig.height = 7, eval = FALSE}
display.brewer.all()
```

Match each number with its corresponding interval to be able to decide its colour.
```{r, eval = FALSE}
col <- colors[findInterval(values_ine, my_pallete,
                           all.inside=TRUE)] # Setting the final colors
```

In case we want to plot the population of certain municipalities in a province, we must use a sequential pallete.

```{r, eval = FALSE}
pallete_colour <- "OrRd" # Scale of oranges and reds

```


Let's say that `n` is the number of colour intervals.
```{r, eval = FALSE}
n <- 5
```


The brewer.pal function builds a pallete with `n` intervals and the `pallete_colour` colour.
```{r, eval = FALSE}
values_ine <- shp_merged@data[[value]] # Values we want to plot are stored in the shape@data data frame
colors <- brewer.pal(n, pallete_colour) # A pallete from RColorBrewer 

```


The style is the distribution of the colour within data. The `classIntervals` function generates numerical intervals. The upper and lower limits of these intervals are named breaks.
```{r, eval = FALSE}
style <- "quantile" 
brks <- classIntervals(values_ine, n = n, style = style)
my_pallete <- brks$brks # my pallete contains the breaks
```


Plot the map and set title and legends.

```{r, fig.width = 8, fig.height = 7, eval = FALSE}

raster::plot(shp_merged,col = col) # Plot the map
title_plot <- "Población total por municipios en La Rioja"

title(main = title_plot)
legend(legend = leglabs(round(my_pallete)), fill = colors,x = "bottomright")
```




#### Other examples

Let's try to plot data at a different administrative level. The next dataset collects the spanish population for all the provinces in a specific range of years. You can find the link in the README.MD file as well. [Link here](http://www.ine.es/jaxiT3/Tabla.htm?t=2852&L=0)

The process is almost the same as the previous one.

```{r, eval = FALSE}
df <- as.data.frame(read.px("/home/ncarvalho/Descargas/2852.px"))
names(df) # List the column names  of the data frame
```


```{r, eval = FALSE}
df[[by]] <- sapply(df$Provincias,
                   function(x) strsplit(x = as.character(x), split = " ")[[1]][1])
```

```{r, eval = FALSE}
df$Periodo <- as.numeric(as.character(df$Periodo))
```

```{r, eval = FALSE}
df <- df[df$Sex == "Total" & df$Periodo == year,]
```

Remember that first we have to create the shapefile. We can't use the previous shapefile provided that the level is not the same.

```{r, eval = FALSE}
level <- "Provincias"
canarias <- FALSE
year <- 2016
scale <- "6m"
```

Generate the map again.  

```{r, message = FALSE, eval = FALSE}
shp <- siane_map(obj = obj, canarias = canarias, year = year, level = level, scale = scale)
```

```{r, eval = FALSE}
value <- "value"
by <- "codes"
```


```{r, eval = FALSE}
shp_merged <- siane_merge(shp = shp, df = df, by = by, level = level, value = value)
```


#### Plot the map.


```{r, fig.width = 7, fig.height = 7, eval = FALSE}
pallete_colour <- "BuPu"
n <- 7
style <- "kmeans"


values_ine <- shp_merged@data[[value]] # Values we want to plot are stored in the shape@data data frame
colors <- brewer.pal(n, pallete_colour) # A pallete from RColorBrewer 


brks <- classIntervals(values_ine, n = n, style = style)
my_pallete <- brks$brks


col <- colors[findInterval(values_ine, my_pallete,
                           all.inside=TRUE)] # Setting the final colors


raster::plot(shp_merged,col = col) # Plot the map

title_plot <- "Población de España a nivel de provincias"

title(main = title_plot)
legend(legend = leglabs(round(my_pallete)), fill = colors,x = "bottomright")




```


## Combining Siane with other libraries

Please download this data [link](http://www.ine.es/jaxi/Tabla.htm?path=/t20/e245/p05/a2002/l0/&file=00008001.px&L=0). This dataset contains the number of inhabitants of Barcelona's municipalities per five year age groups.
We are now trying to combine leaflet library with Siane to make a higher quality visualization. 
In this visualization I will plot the percentage of girls whose ages are in the range of 0-4 years.

```{r, message = FALSE, warning = FALSE, eval = FALSE}
library(leaflet)
library(data.table)
```

Reading the map
```{r, eval = FALSE}
shp <- siane_map(obj = obj, canarias = FALSE, year = 2016,level = "Municipios" ,scale = "3m")
```

Reading the data
```{r, eval = FALSE}
df <- as.data.frame(read.px("/home/ncarvalho/Descargas/00008001.px"))
df <- as.data.table(df)
```

Calculating the sum of the inhabitants of all ages per sex.
```{r, eval = FALSE}
df_sum <- df[, sum(value),by = c("Municipios", "Sexo") ]
```

Calculating the percentage of each group of age per municipality and gender
```{r, eval = FALSE}
df_withsums <- merge(df, df_sum, by = c("Municipios","Sexo"))
df_withsums$perc <- df_withsums$value/df_withsums$V1*100
```

Filtering the data: Women with 0-4 years old.

```{r, eval = FALSE}

df_withsums <- df_withsums[df_withsums$Sexo=="Mujeres"&
                           df_withsums$Edad..grupos.quinquenales.=="0-4"]
```


Extracting the territory code for all the municipalities
```{r, eval = FALSE}
by <- "codes"
value <- "perc"

df_withsums[[by]] <- sapply(df_withsums$Municipios,
                   function(x) strsplit(x = as.character(x), split = " ")[[1]][1])

```


Merging spatial data and statistical data.

```{r, eval = FALSE}
shp_merged <- siane_merge(shp = shp, df = df_withsums, by = by,value = value, level = "Municipios")
```

Plot the data
```{r , fig.width = 7, fig.height = 7, eval = FALSE, message = FALSE, warning = FALSE}
pal <- colorNumeric(palette = "YlOrRd", domain = shp_merged@data$perc)

leaflet(shp_merged) %>%  # 
  addTiles() %>%    # Add the world map 
  addPolygons(color = "#444444", weight = 1, smoothFactor = 0.5,  # options
    opacity = 1.0, fillOpacity = 0.5, # options
    fillColor = ~colorQuantile("YlOrRd", perc)(perc), # Choose the scale, Choose the column to plot
    highlightOptions = highlightOptions(color = "white", weight = 2, # options
      bringToFront = TRUE)) %>%# options

  addLegend("bottomright", pal = pal, values = shp_merged$perc , title = "Porcentaje de mujeres con 0-4 anyos", opacity = 1)



```

This leaflet documentation is available [here](https://rstudio.github.io/leaflet/shapes.html)




